State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \beta \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Lh\left( t \right) = \delta \left( t \right),
\end{equation}
\todo{Is the above right? I can't think of any way to make $\frac{d}{dt} h(t) + \alpha h(t)$ equal to $\delta(t)$}
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \alpha v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\lambda = 1-\alpha T$ and is an iid spatially colored, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\varepsilon_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

@@

\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int\limits_{\Omega}\int\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 

>>>>

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \beta \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Dh\left( t \right) = \delta \left( t \right),
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \alpha v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\xi = 1-\zeta T$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

	&=& Q\left(\mathbf{x}_t,u_t(\mathbf{r})\right) + \mathbf{e}_t \\

@@

\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
